\documentclass[11pt,reqno]{amsart}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amsthm}
\usepackage{mathtools}
\usepackage{latexsym}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{float}
\usepackage{graphbox}
\usepackage{textcomp}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue,citecolor=red]{hyperref}
\usepackage{array}
\usepackage{comment}
\usepackage{appendix}
\usepackage{enumitem}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{question}[theorem]{Question}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left\lvert#1\right\rvert}
\def\R{\mathbb{R}}
\def\C{\mathbb{C}}
\def\N{\mathbb{N}}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Prob}{P}
\DeclareMathOperator{\diag}{diag}

% Draft provenance colors.
% Uncolored text is pre-existing draft material.
% `revderived`: assistant organization/paraphrase of pre-existing draft text.
% `revverified`: prose translated from verified proof-route evidence.
% `revnewmath`: genuinely new mathematical exposition added in this pass.
\definecolor{RevisionDerived}{RGB}{105,105,105}
\definecolor{RevisionVerified}{RGB}{0,105,120}
\definecolor{RevisionNewMath}{RGB}{0,45,150}
\newenvironment{revderived}{\begingroup\color{RevisionDerived}}{\endgroup}
\newenvironment{revverified}{\begingroup\color{RevisionVerified}}{\endgroup}
\newenvironment{revnewmath}{\begingroup\color{RevisionNewMath}}{\endgroup}
% Backwards-compatible aliases used by the included appendix body.
\newenvironment{revlatest}{\begin{revnewmath}}{\end{revnewmath}}
\newenvironment{revrecent}{\begin{revverified}}{\end{revverified}}
\newenvironment{revearlier}{\begin{revderived}}{\end{revderived}}

\title[Sharp upper tails for PPT moments]{Sharp upper tails for partial-transpose moments on the Hilbert--Schmidt sphere}
\author{}
\date{}

\begin{document}
\begin{revverified}
\begin{abstract}
We isolate the upper-bound half of a sharp large-deviation principle for the
\(k\)-th partial-transpose moment of a random induced state in the balanced
regime \(s(d)\sim \lambda d^2\).  The upper-tail speed is
\(d^{2+2/k}\), and the rate predicted by the one-column spike mechanism is
\(\lambda\varepsilon^{1/k}\).  The proof developed here explains why no
larger upper-tail probability can occur: after removing a background typical
set, the bad event is forced outside a spherical neighbourhood whose
isoperimetric cost is exactly
\(\exp\{-\lambda\varepsilon^{1/k}d^{2+2/k}+o(d^{2+2/k})\}\).
The article is organized as a conditional proof route; the older
\(d^2\)-scale concentration appendix is retained as a supplier for the
typical-set estimate, not as the sharp theorem.
\end{abstract}
\end{revverified}

\maketitle

\begin{revnewmath}
\begin{remark}[Draft provenance colors]
The colors are provenance labels, not decoration.  Uncolored text is inherited
from the pre-existing draft.  Gray text is assistant organization or paraphrase
of that draft.  Teal text is translated from proof-route statements.  Dark blue text is new mathematical exposition added in
this pass.  The color environments are local and can be redefined to no-ops
for a submission version.
\end{remark}
\end{revnewmath}

\begin{revnewmath}
\section{Scope and relation with the PPT threshold}
\label{sec:scope-aubrun}

Aubrun's theorem on partial transposition of random induced states is a
quantitative threshold result.  In particular, the statement is not merely
that the balanced PPT threshold is the constant \(4\): the theorem also gives
high-probability estimates, with explicit inequalities controlling the two
sides of the threshold.  The present paper should therefore not be read as a
replacement proof of Aubrun's threshold theorem unless those probability
bounds are restated and matched in the same normalization.

The target here is narrower.  We study a fixed moment order \(k\ge3\), the
balanced environment \(s(d)\sim\lambda d^2\), and the logarithmic upper tail
of the partial-transpose moment at speed
\[
  v_{k,d}=d^{2+2/k}.
\]
The mechanism is a large-deviation mechanism for a moment functional: a spike
of spherical radius \(d^{-1+1/k}\) creates a moment excess of order one, and
the ambient Hilbert--Schmidt sphere charges such a displacement with cost
\(\exp\{-\lambda a\,d^{2+2/k}+o(d^{2+2/k})\}\).  This is compatible with the
PPT-threshold picture, but it is a different assertion from Aubrun's
quantified threshold bounds.
\end{revnewmath}

\begin{revverified}
\section{Sharp logarithmic upper tail}
\label{sec:sharp-log-upper-tail}

Fix an integer \(k\ge3\), a level \(\varepsilon>0\), and a balanced
environment sequence \(s(d)\) with \(s(d)/d^2\to\lambda>0\).  Let \(X_d\) be
uniform on the Hilbert--Schmidt unit sphere in \(M_{d^2,s(d)}(\C)\), and let
\[
  p_d^\uparrow(\varepsilon)
  =
  \Prob\left\{
    \left|
      m_{k,d}(X_d)-\E m_{k,d}(X_d)
    \right|\ge \varepsilon
  \right\}
\]
denote the absolute-deviation upper-tail probability for the normalized
background moment used in the formal Appendix B development.  The speed and
candidate rate are
\[
  v_{k,d}=d^{2+2/k},
  \qquad
  I_k(\varepsilon)=\lambda\varepsilon^{1/k}.
\]

\begin{theorem}[Formalized sharp upper endpoint]
\label{thm:sharp-upper-endpoint}
Assume the following inputs for the concrete spherical matrix model:
\begin{enumerate}[leftmargin=*]
\item the full spherical isoperimetric inequality in the ambient real
  dimension \(2d^2s(d)\);
\item the dimension identifications between the matrix model and that real
  sphere;
\item eventual positive mass of the one-sided favourable upper-tail event;
\item eventual half-mass of the background typical set, obtained from the
  moment and operator-norm bad-set estimates;
\item the deterministic mixed-word expansion bound in the sharp spherical
  neighbourhood, together with the corresponding scalar term limits.
\end{enumerate}
Then for every \(\eta>0\), eventually in \(d\),
\[
  \frac{\log p_d^\uparrow(\varepsilon)}{v_{k,d}}
  \le
  -I_k(\varepsilon)+\eta .
\]
\end{theorem}
\end{revverified}

\begin{revnewmath}
\begin{theorem}[Main upper large-deviation bound]
\label{thm:main-upper-large-deviation}
Let \(k\ge3\), \(\varepsilon>0\), and \(s(d)/d^2\to\lambda>0\).  Suppose that
the five input estimates listed in Theorem~\ref{thm:sharp-upper-endpoint}
hold for the balanced spherical model.  Then
\[
  \limsup_{d\to\infty}
  \frac{1}{d^{2+2/k}}
  \log
  \Prob\left\{
    \left|
      m_{k,d}(X_d)-\E m_{k,d}(X_d)
    \right|\ge \varepsilon
  \right\}
  \le
  -\lambda\varepsilon^{1/k}.
\]
\end{theorem}

\begin{proof}
Apply Theorem~\ref{thm:sharp-upper-endpoint} with an arbitrary
\(\eta>0\), take the \(\limsup\), and then let \(\eta\downarrow0\).  The
content is therefore the formalized endpoint itself: all dimensional
normalizations, local expansion estimates, and spherical-isoperimetric
constants have already been reduced to the concrete model.
\end{proof}

\begin{proof}[Proof route]
Choose a slack parameter and set the spherical radius so that its \(k\)-th
power is just below \(\varepsilon\).  On the background typical set, the local
expansion of the \(k\)-th moment separates into the pure radial spike term,
the background term, and mixed words.  The deterministic mixed-word estimate
and its scalar limits make the mixed contribution smaller than the available
slack, so every point in the upper-tail event must lie outside the sharp
spherical neighbourhood of the typical set.

The typical set has probability at least \(1/2\).  Spherical isoperimetry
therefore bounds the probability of the complement of its radius-\(r_d\)
neighbourhood by
\[
  \exp\left(-\frac{(2d^2s(d)-1)r_d^2}{2}\right).
\]
The radius is chosen at the spike scale, and the scalar identities reduce the
exponent to
\[
  -\lambda\,a + o(1)
\]
after division by \(d^{2+2/k}\), where \(a^k\) is the admissible moment
increase.  Letting the slack tend to zero gives \(a\to\varepsilon^{1/k}\),
which yields the claimed \(-\lambda\varepsilon^{1/k}\) upper rate.
\end{proof}

\begin{lemma}[Radius-to-rate computation]
\label{lem:radius-to-rate}
Let \(N_d=d^2\), let \(s(d)/N_d\to\lambda\), and let
\(v_{k,d}=d^{2+2/k}\).  For a fixed \(a\ge0\), set
\[
  r_d^2=\frac{a\,v_{k,d}}{N_d^2},
  \qquad
  n_d=2N_ds(d).
\]
Then
\[
  \frac{(n_d-1)r_d^2}{2v_{k,d}}\longrightarrow \lambda a .
\]
\end{lemma}

\begin{proof}
Since \(r_d^2=a v_{k,d}/N_d^2\),
\[
  \frac{(n_d-1)r_d^2}{2v_{k,d}}
  =
  a\left(\frac{s(d)}{N_d}-\frac{1}{2N_d^2}\right).
\]
The first term tends to \(\lambda\), and the second tends to \(0\).
\end{proof}

\begin{theorem}[Sharp exponent, conditional on the matching lower spike]
\label{thm:sharp-optimal-exponent}
Assume, in addition to the hypotheses of
Theorem~\ref{thm:sharp-upper-endpoint}, the one-column lower spike
construction: for every \(\eta>0\), eventually in \(d\),
\[
  -I_k(\varepsilon)-\eta
  \le
  \frac{\log p_d^\uparrow(\varepsilon)}{v_{k,d}} .
\]
Then
\[
  \lim_{d\to\infty}
  \frac{\log p_d^\uparrow(\varepsilon)}{d^{2+2/k}}
  =
  -\lambda\varepsilon^{1/k}.
\]
\end{theorem}

\begin{proof}
The lower spike gives
\(\liminf_d v_{k,d}^{-1}\log p_d^\uparrow(\varepsilon)
\ge -I_k(\varepsilon)\).  The sharp upper endpoint gives
\(\limsup_d v_{k,d}^{-1}\log p_d^\uparrow(\varepsilon)
\le -I_k(\varepsilon)\).  The two inequalities force the limit and identify
the optimal exponent.
\end{proof}

\begin{remark}[What is stronger than the concentration appendix]
The concentration theorem later in this file controls ordinary fluctuations
around the mean on an exponential scale supplied by the ambient sphere.  The
sharp endpoint above is a large-deviation statement at the smaller spike
speed \(d^{2+2/k}\), and it identifies the exact constant in the exponent.
Thus the concentration argument is only one supplier for the typical
background set; it is not the final optimality theorem.
\end{remark}
\end{revnewmath}

\begin{revverified}
\section{Proof bookkeeping}
\label{sec:proof-bookkeeping}

The proof separates the probabilistic argument into an upper inequality and a
matching one-column lower construction.  Both sides are normalized at the same
speed and rate:
\[
  v_d=d^{2+2/k},
  \qquad
  I(\varepsilon)=\lambda\varepsilon^{1/k}.
\]
The bookkeeping point is that the upper argument is an optimal-exponent
argument, not merely an ordinary concentration estimate: the concentration
appendix supplies typical background control, while the sharp speed comes from
the spike displacement and the cap cost on the Hilbert--Schmidt sphere.
\end{revverified}

\begin{revderived}
\section{Introduction}

Let \(X\) be uniformly distributed on the Hilbert--Schmidt unit sphere in
\(M_{d^2,s}(\C)\).  The matrix \(XX^*\) is the normalized induced random
state obtained from a Gaussian matrix after radial projection, and
\((XX^*)^\Gamma\) denotes its partial transpose with respect to the tensor
decomposition \(M_{d^2}(\C)=M_d(\C)\otimes M_d(\C)\).  The quantities studied
in this article are the moments
\[
  f_k(X)=\Tr\left(((XX^*)^\Gamma)^k\right),
  \qquad k\ge1.
\]

The balanced regime is \(s\asymp d^2\).  This is the natural threshold scale
for partial transposition of random induced states: the matrix dimension is
\(d^2\), the number of environmental columns is comparable to \(d^2\), and the
real dimension of the Hilbert--Schmidt sphere is
\[
  2d^2s-1\asymp d^4.
\]
This large ambient dimension is the source of the concentration estimate
below.

There is, however, a useful obstruction.  The map \(X\mapsto f_k(X)\) is not
treated as globally Lipschitz at the optimal scale.  A global Lipschitz bound
would be too crude, because it would have to control rare configurations with
large operator norm.  Instead, the proof constructs a high-probability set
\(\Omega\) on which the relevant operator norms are of order \(d^{-2}\).  On
this set the trace-power perturbation estimate gives the correct local
Lipschitz constant.  A localized form of Levy's lemma then converts the local
Lipschitz estimate and the smallness of \(\Omega^c\) into concentration on the
whole sphere.

\section{Model and main result}

The ambient space is the Hilbert--Schmidt unit sphere
\[
  S_{HS}(d^2,s)
  =
  \{X\in M_{d^2,s}(\C):\norm X_2=1\},
\]
with its uniform surface probability measure.  The observable is
\[
  f_k(X)=\Tr\left(((XX^*)^\Gamma)^k\right),
\]
where \(\Gamma\) denotes partial transposition on the second tensor factor of
\(M_d(\C)\otimes M_d(\C)\).

\begin{theorem}[Main concentration estimate]
\label{thm:main-concentration}
Fix \(k\ge1\) and \(\lambda>0\).  There are constants \(c,C>0\), depending
only on \(k\) and \(\lambda\), such that the following holds for all
sufficiently large \(d\).  If
\[
  \frac{\lambda}{2}d^2\le s\le 2\lambda d^2
\]
and \(X\) is uniform on \(S_{HS}(d^2,s)\), then for every \(\varepsilon>0\),
\[
  \Prob\left\{
    \left|
      f_k(X)-\E f_k(X)
    \right|
    \ge
    \varepsilon d^{-(2k-2)}
  \right\}
  \le
  C e^{-c d^2}
  +
  C\exp\left(-c\frac{d^4\varepsilon^2}{k^2}\right).
\]
\end{theorem}

The first term in the bound is the cost of discarding the bad set where the
operator-norm controls may fail.  The second term is the spherical
concentration term.  The exponent \(d^4\) is the real dimension of the sphere,
while the factor \(k^2\) comes from the local Lipschitz constant of the
degree-\(k\) trace polynomial.

\begin{remark}
For fixed \(\varepsilon>0\), the second term is much smaller than
\(\exp(-c d^2)\).  We keep the two terms separated because the proof naturally
produces them from different mechanisms: probabilistic construction of the
good set and spherical concentration on the good set.
\end{remark}

\section{Proof architecture}

The proof of Theorem~\ref{thm:main-concentration} has four conceptual steps.

\begin{enumerate}[leftmargin=*]
\item \emph{Localize the concentration inequality.}
  Levy concentration is applied not directly to \(f_k\), but to a Lipschitz
  extension of \(f_k\) from a high-probability good set.  This produces a tail
  estimate with two errors: the spherical Levy tail and the probability of the
  bad set.

\item \emph{Build the good set.}
  On the good set, the operator norm of \((XX^*)^\Gamma\) and the relevant
  diagonal gamma quantities are bounded at the correct \(d^{-2}\) scale.  The
  complement of this good set is exponentially small.

\item \emph{Prove the deterministic Lipschitz bound on the good set.}
  The trace-power telescoping identity reduces the variation of \(f_k\) to a
  trace-norm perturbation bound.  Hilbert--Schmidt ideal inequalities and the
  operator-norm control on the good set yield the desired local Lipschitz
  constant.

\item \emph{Replace the median by the expectation.}
  The localized concentration estimate first gives concentration around a
  median.  Since \(f_k\) is uniformly bounded, the tail estimate can be
  integrated to show that the median and the mean differ by a smaller error.
  This gives the final concentration inequality around \(\E f_k\).
\end{enumerate}

\subsection{Why the local Lipschitz constant has the right scale}

The deterministic perturbation bound starts from
\[
  \abs{\Tr(A^k)-\Tr(B^k)}
  \le
  k\max(\norm A_\infty,\norm B_\infty)^{k-1}\norm{A-B}_1.
\]
Applied to
\[
  A=(XX^*)^\Gamma,\qquad B=(YY^*)^\Gamma,
\]
the trace-norm difference is controlled by
\[
  \norm{A-B}_1
  \lesssim
  d\,\max(\norm X_\infty,\norm Y_\infty)\norm{X-Y}_2.
\]
On the good set one has
\[
  \norm{(XX^*)^\Gamma}_\infty \lesssim d^{-2},
  \qquad
  \norm X_\infty \lesssim d^{-1}.
\]
Consequently
\[
  \abs{f_k(X)-f_k(Y)}
  \lesssim
  k d^{-2(k-1)}\norm{X-Y}_2
  \qquad (X,Y\in\Omega).
\]
Thus the relevant Lipschitz constant is
\[
  L_{k,d}\asymp k d^{-2k+2}.
\]
Since the sphere has real dimension \(n\asymp d^4\), Levy's lemma gives
\[
  \exp\left(-c\,\frac{n t^2}{L_{k,d}^2}\right)
  =
  \exp\left(-c\,\frac{d^4\varepsilon^2}{k^2}\right)
\]
when \(t=\varepsilon d^{-(2k-2)}\).  This is the scale appearing in
Theorem~\ref{thm:main-concentration}.

\subsection{The good set}

The good set \(\Omega\) is the intersection of two events.  The first controls
the operator norm of \(X\), and follows from standard Gaussian operator-norm
estimates after writing \(X=G/\norm G_2\).  The second controls
\((XX^*)^\Gamma\).  For this, one decomposes the partially transposed Wishart
matrix into its diagonal and off-diagonal parts.  The diagonal part is
controlled by gamma concentration, while the off-diagonal part is controlled
by Aubrun's moment estimate and Markov's inequality applied to a high even
moment.

The output is the estimate
\[
  \Prob(\Omega^c)\le C e^{-c d^2}.
\]
This probability is not negligible compared with the final bound; it is
exactly the first term in Theorem~\ref{thm:main-concentration}.

\subsection{Median and expectation}

The localized Levy argument first gives concentration around a median.  To
pass to the expectation, we use the uniform bound \(\abs{f_k(X)}\le1\) for
\(k\ge2\), while \(f_1(X)=1\).  Integrating the median tail only up to the
bounded range gives
\[
  \abs{\E f_k-M_k}
  \le
  C e^{-c d^2}+C_{k,\lambda}k d^{-2k}.
\]
This error is smaller than the target scale
\(\varepsilon d^{-(2k-2)}\) except in the regime where the stated probability
bound is already trivial after increasing the constant \(C\).  This completes
the reduction from median concentration to mean concentration.

\subsection{Detailed proof}

The rest of the article gives the detailed estimates in the order just
described.
\end{revderived}

\appendix
\input{AppendixB_article_replacement.tex}

% Standalone-only bibliography for compiling this wrapper outside the article.
% Do not paste this block into a host article that already uses a bibliography.
\begin{thebibliography}{2}
\bibitem{Aubrun12}
G.~Aubrun,
\emph{Partial transposition of random states and non-centered semicircular
distributions},
Random Matrices: Theory Appl. \textbf{1} (2012), 1250001.

\bibitem{AubrunSzarek2017}
G.~Aubrun and S.~J. Szarek,
\emph{Alice and Bob Meet Banach: The Interface of Asymptotic Geometric
Analysis and Quantum Information Theory},
Mathematical Surveys and Monographs, vol.~223, American Mathematical
Society, 2017.
\end{thebibliography}

\end{document}
